---
title: "Summit_cpue_samp_size"
output: html_document
date: "2026-03-16"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
library(readxl)
library(knitr)
library(tidyverse)
library(janitor)
library(nleqslv)
source("cpue.R")
```


```{r}
# Install nleqslv if you don't have it
# install.packages("nleqslv")
# Parameters
n1 <- 190              
mean_cpue <- 0.1222710       
var_mean_cpue <- 0.0383483^2  # example variance, replace with your var_cpue
alpha = 0.15
beta = 0.3
z_alpha <- qnorm(1-alpha/2)       
z_beta <- qnorm(beta)          
zfunc <- z_alpha - z_beta 


compute_n2_nleqslv <- function(d, zfunc) 
{
  
  zfunc = zfunc
  f_n2_eq <- function(n2) {
    denom <- var_mean_cpue + (n1 / n2) * (var_mean_cpue / mean_cpue^2) * (mean_cpue + d)^2
    d^2 / denom - zfunc^2
  }
  
  solution <- nleqslv(x = 100, fn = f_n2_eq, control = list(allowSingular = TRUE))
  
  #check for termination, skip if terminated
  if (solution$termcd != 1)
    {
      return(NA)
    } 
      else {
      return(solution$x)
  }
}



```

As we can see above, the parameters derived from the CPUE bootstrapping of 1990 are: sample size = 190, mean cpue= 0.12, var in mean cpue = 0.0014. Our starting values of alpha and beta correspond to a = 0.15 and b = 0.7.

I will proceed with this estimate as a standard of 0.3 units of change as that was the reference in the paper. 


Let's start with this very low confidence estimtae:


```{r, echo = FALSE}

print("sample size for a change of 0.3 units cpue with alpha = 0.15 and power = 0.7")
print(compute_n2_nleqslv(0.3, zfunc = zfunc))


d_values <- seq(0.15, 0.5, by = 0.01)

# Compute required n2 for each d
n2_values <- sapply(d_values, compute_n2_nleqslv, zfunc = zfunc)

# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]


df <- data.frame(d = d_values_valid, n2 = n2_values_valid)

# ggplot
ggplot(df, aes(x = d, y = n2)) +
  geom_point(color = "darkgreen") +
  geom_line(color = "darkgreen") +
  labs(x = "Absolute Difference in CPUE", y = "Number of Traps", title = "alpha = 0.15, power = 0.7") +
  theme_minimal()


```

As we increase our accuracy and precision, we will see an increase in sample size requirements. 

```{r, echo = FALSE}


alpha = 0.15
beta = 0.2
z_alpha <- qnorm(1-alpha/2)       
z_beta <- qnorm(beta)  
zfunc <- z_alpha- z_beta  
print("sample size for a change of 0.3 units cpue and power of 0.8")
print(compute_n2_nleqslv(0.3, zfunc = zfunc))
n2_values <- sapply(d_values, compute_n2_nleqslv,zfunc = zfunc)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot


df <- data.frame(d = d_values_valid, n2 = n2_values_valid)

# ggplot
ggplot(df, aes(x = d, y = n2)) +
  geom_point(color = "darkgreen") +
  geom_line(color = "darkgreen") +
  labs(x = "Absolute Difference in CPUE", y = "Number of Traps", title = "alpha = 0.15, power = 0.8") +
  theme_minimal()


alpha = 0.15
beta = 0.15
z_alpha <- qnorm(1-alpha/2)       
z_beta <- qnorm(beta)  
zfunc <- z_alpha- z_beta  
print("sample size for a change of 0.3 units cpue and power of 0.85")
print(compute_n2_nleqslv(0.3, zfunc = zfunc))
n2_values <- sapply(d_values, compute_n2_nleqslv,zfunc = zfunc)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]


df <- data.frame(d = d_values_valid, n2 = n2_values_valid)

# ggplot
p <- ggplot(df, aes(x = d, y = n2)) +
  geom_point(color = "darkgreen") +
  geom_line(color = "darkgreen") +
  labs(x = "Absolute Difference in CPUE", y = "Number of Traps",  title = "alpha = 0.15, power = 0.85") +
  theme_minimal()

# Save

alpha = 0.1
beta = 0.2
z_alpha <- qnorm(1-alpha/2)       
z_beta <- qnorm(beta)  
zfunc <- z_alpha- z_beta  
print("sample size for a change of 0.3 units cpue and power of 0.8")
print(compute_n2_nleqslv(0.2, zfunc = zfunc))
n2_values <- sapply(d_values, compute_n2_nleqslv,zfunc = zfunc)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot

df <- data.frame(d = d_values_valid, n2 = n2_values_valid)

# ggplot
p <- ggplot(df, aes(x = d, y = n2)) +
  geom_point(color = "darkgreen") +
  geom_line(color = "darkgreen") +
  labs(x = "Absolute Difference in CPUE", y = "Number of Traps",  title = "alpha = 0.1, power = 0.8") +
  theme_minimal()

ggsave("PLOT_TO_USE.png", dpi = 300)
###THIS IS THE ONE WE SHOULD SAVE!
# Save


```



Generating a dataframe which I will save separately for the sample size.
```{r}


get_zalpha<- function(alpha, beta)
{
  return(qnorm(1-alpha/2)-qnorm(beta))
}

list_alpha<- c(0.05, 0.1,0.15)
list_beta<- c(0.1,0.15,0.2,0.25,0.3)
#power is 1-beta
#list_zalpha<- sapply(list_alpha, get_zalpha)          

#list_beta<- c(0.7,0.75,0.8,0.85,0.9,0.95)
list_d<- c(0.01,0.05,0.1,0.15,0.2, 0.25,0.3)


df_combos <- expand.grid(
  alpha = list_alpha,
  beta = list_beta,
  d = list_d
)
df_combos<- df_combos%>%
  rowwise()%>%
  mutate(z_func= get_zalpha(alpha, beta))%>%
  ungroup()



get_zalpha(0.1, 0.2)
compute_n2_nleqslv(d = 0.2, zfunc = get_zalpha(0.1, 0.2))
df_combos

 df_combos <- df_combos %>%
  rowwise() %>%  # ensures mutate sees each row individually
  mutate(result = compute_n2_nleqslv(d = d, zfunc = z_func)) %>%
  ungroup()%>%
  select(-c(z_func))%>%
  mutate(
    power = 1-beta
  )%>%
  select(-c(beta))
 
 acceptable_ranges<- df_combos%>%
   filter(result < 450)%>%
   filter(power == c(0.75,0.8, 0.85))

 
write.csv(acceptable_ranges, "80_power.csv")
write.csv(df_combos,"samp_size_by_precision_summit.csv")

```