---
title: "Summit_cpue_samp_size"
output: html_document
date: "2026-03-16"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
library(readxl)
library(knitr)
library(tidyverse)
library(janitor)
library(nleqslv)
source("cpue.R")
```


```{r}
# Install nleqslv if you don't have it
# install.packages("nleqslv")
# Parameters
n1 <- 190              
mean_cpue <- 0.1222710       
var_mean_cpue <- 0.0383483^2  # example variance, replace with your var_cpue
z_alpha <- 1.44        
z_beta <- 0.7          
zfunc <- z_alpha + z_beta 


compute_n2_nleqslv <- function(d) {
  f_n2_eq <- function(n2) {
    denom <- var_mean_cpue + (n1 / n2) * (var_mean_cpue / mean_cpue^2) * (mean_cpue + d)^2
    d^2 / denom - zfunc^2
  }
  
  solution <- nleqslv(x = 100, fn = f_n2_eq, control = list(allowSingular = TRUE))
  
  #check for termination, skip if terminated
  if (solution$termcd != 1)
    {
      return(NA)
    } 
      else {
      return(solution$x)
  }
}


```

As we can see above, the parameters derived from the CPUE bootstrapping of 1990 are: sample size = 190, mean cpue= 0.12, var in mean cpue = 0.0014. Our starting values of alpha and beta correspond to a = 0.15 and b = 0.7.

I will proceed with this estimate as a standard of 0.3 units of change as that was the reference in the paper. 


Let's start with this very low confidence estimtae:


```{r, echo = FALSE}

print("sample size for a change of 0.3 units cpue")
print(compute_n2_nleqslv(0.3))


d_values <- seq(0.15, 0.5, by = 0.01)

# Compute required n2 for each d
n2_values <- sapply(d_values, compute_n2_nleqslv)

# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot
plot(d_values_valid, n2_values_valid, type = "b", pch = 19, col = "darkgreen",
     xlab = "change (d, units cpue)", ylab = "Required n2",
     main = "Required Sample Size n2 for alpha = 0.15, beta = 0.7",
     ylim = c(min(n2_values_valid), max(n2_values_valid)))

# Vector of effect sizes

```

As we increase our accuracy and precision, we will see an increase in sample size requirements. 

```{r, echo = FALSE}


z_alpha <- qnorm(1-0.15/2)        
z_beta <- 0.8
zfunc <- z_alpha + z_beta  
print("sample size for a change of 0.3 units cpue")
print(compute_n2_nleqslv(0.3))
n2_values <- sapply(d_values, compute_n2_nleqslv)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot

plot(d_values_valid, n2_values_valid, type = "b", pch = 19, col = "darkgreen",
     xlab = "change (d, units cpue)", ylab = "Required n2",
     main = "Required Sample Size n2 for alpha = 0.15, beta = 0.8",
     ylim = c(min(n2_values_valid), max(n2_values_valid)))


z_alpha <- qnorm(1-0.1/2)        
z_beta <- 0.8
zfunc <- z_alpha + z_beta  
print("sample size for a change of 0.3 units cpue")
print(compute_n2_nleqslv(0.3))
n2_values <- sapply(d_values, compute_n2_nleqslv)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot

plot(d_values_valid, n2_values_valid, type = "b", pch = 19, col = "darkgreen",
     xlab = "change (d, units cpue)", ylab = "Required n2",
     main = "Required Sample Size n2 for alpha = 0.1, beta = 0.8",
     ylim = c(min(n2_values_valid), max(n2_values_valid)))




z_alpha <- qnorm(1-0.1/2)        
z_beta <- 0.85
zfunc <- z_alpha + z_beta  
print("sample size for a change of 0.3 units cpue")
print(compute_n2_nleqslv(0.3))
n2_values <- sapply(d_values, compute_n2_nleqslv)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot

plot(d_values_valid, n2_values_valid, type = "b", pch = 19, col = "darkgreen",
     xlab = "change (d, units cpue)", ylab = "Required n2",
     main = "Required Sample Size n2 for alpha = 0.1, beta = 0.85",
     ylim = c(min(n2_values_valid), max(n2_values_valid)))


z_alpha <- qnorm(1-0.05/2)        
z_beta <- 0.85
zfunc <- z_alpha + z_beta  
print("sample size for a change of 0.3 units cpue")
print(compute_n2_nleqslv(0.3))
n2_values <- sapply(d_values, compute_n2_nleqslv)
# Filter valid results
valid_idx <- !is.na(n2_values)
d_values_valid <- d_values[valid_idx]
n2_values_valid <- n2_values[valid_idx]

# Plot

plot(d_values_valid, n2_values_valid, type = "b", pch = 19, col = "darkgreen",
     xlab = "change (d, units cpue)", ylab = "Required n2",
     main = "Required Sample Size n2 for alpha = 0.05, beta = 0.85",
     ylim = c(min(n2_values_valid), max(n2_values_valid)))

```
